---
title: "NZ SRW Calving interval Report"
author: "Anthony Davidson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  pdf_document: default
  html_document: default
geometry: margin=1in
fontsize: 14pt
---

#Raw data
<!--The data frame imported from Will is as follows:-->

```{r setup, echo=FALSE, warning = FALSE, message = FALSE}
#NZJFMS manuscript
#15032017
#Anthony Davidson

#packages
library(boot) 
library(dplyr)
library(ggplot2)
library(qpcR)

#set working dir
#laptop
setwd("C:/Users/s435389/Dropbox/NZJMFR Publication Calving interval/NZJMFR Calving Analysis/Final code for publication")
#desktop
#setwd("C:/Users/s435389/Dropbox (Population_Stats)/NZJMFR Publication Calving interval/NZJMFR Calving Analysis/Final code for publication")

###############################DATA####################################
Data<- read.csv("RawCI.csv", header=T, quote="\"")

#Structure of data set
#str(Data)

#year recorded
Year<-unique(Data$Calves.1)

#calving interval observed in 2010
year2010a<-c(3,3,2)
year2010 <- filter(Data,Calves.1 < 2011)
year2010 <- year2010$Interval.1[!is.na(year2010$Interval.1)]

#calving interval observed in 2011
year2011a<-c(3,3,2,3,3,3,3,3,3,3,3,3,3,3,2)
year2011 <- filter(Data,Calves.1 < 2012)
year2011 <- year2011$Interval.1[!is.na(year2011$Interval.1)]

#calving interval observed in 2012
year2012a<-c(3,3,2,3,3,3,3,3,3,3,3,3,3,3,2,
            6,4,4,4,4,4,3,3,3,3)
year2012 <- filter(Data,Calves.1 < 2013)
year2012 <- year2012$Interval.1[!is.na(year2012$Interval.1)]

#calving interval observed in 2013
year2013a<-c(3,3,2,3,3,3,3,3,3,3,3,3,3,3,2,
            6,4,4,4,4,4,3,3,3,3,
            6,5,4,4,4,4,4,3,3,3,3,3,3,3,3,3,3,3,2,2)
full <- c(Data$Interval.1,Data$Interval.2)
year2013<- full[!is.na(unlist(full))]
```

The data is then converted into intervals for each of the years 2010, 2011, 2012 and 2013.

```{r basic, echo=FALSE, message=FALSE, warning=FALSE}
#Confidence intervals
#2010
mean2010<-sum(year2010)/length(year2010)
s2010<-sd(year2010)
SE2010<-s2010/(sqrt(length(year2010)))
n2010<-(length(year2010))
low.qt2010 <- mean2010-(qt(0.975,length(year2010))*SE2010)
high.qt2010 <- mean2010+(qt(0.975,length(year2010))*SE2010)

#2011
mean2011<-sum(year2011)/length(year2011)
s2011<-sd(year2011)
SE2011<-s2011/(sqrt(length(year2011)))
n2011<-(length(year2011))
low.qt2011 <- mean2011-(qt(0.975,length(year2011))*SE2011)
high.qt2011 <- mean2011+(qt(0.975,length(year2011))*SE2011)

#2012
mean2012<-sum(year2012)/length(year2012)
s2012<-sd(year2012)
SE2012<-s2012/(sqrt(length(year2012)))
n2012<-(length(year2012))
low.qt2012 <- mean2012-(qt(0.975,length(year2012))*SE2012)
high.qt2012 <- mean2012+(qt(0.975,length(year2012))*SE2012)

#2013
mean2013<-sum(year2013)/length(year2013)
s2013<-sd(year2013)
SE2013<-s2013/(sqrt(length(year2013)))
n2013<-(length(year2013))
low.qt2013 <- mean2013-(qt(0.975,length(year2013))*SE2013)
high.qt2013 <- mean2013+(qt(0.975,length(year2013))*SE2013)

#Makes data frame to plot
n <- c(length(year2010),length(year2011),length(year2012),length(year2013))
mY <- c(mean(year2010),mean(year2011),mean(year2012),mean(year2013))
year <- Year
low.qt <- c(low.qt2010,low.qt2011,low.qt2012,low.qt2013)
high.qt <- c(high.qt2010,high.qt2011,high.qt2012,high.qt2013)
sd <- c(s2010,s2011,s2012,s2013)
sum.dat <- cbind(year,n,mY,low.qt,high.qt,sd)
sum.dat <-  as.data.frame(sum.dat)

```

A table of the raw results as the year increase

```{r raw table, echo=FALSE}
library(knitr)
kable(sum.dat, format = "markdown")
```


##Raw data plots
The raw intervals can be presented as different means and confidence intervals as the length of the study increases:

```{r raw graph, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#plot data
ggplot(sum.dat, aes(y = mY, x = year)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = low.qt, ymax = high.qt), width = 0.1) +
  theme_bw()
```

We can also plot these as bar charts that show this more obviously:

```{r raw graph 2, eval=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE, include=FALSE}

#PLOTS
par(mfrow=c(2,2))

plot(factor(year2010),xlim=c(0,6),ylim=c(0,40))
title(main="a)",sub="Sample size 3", ylab="Frequency",xlab="Calving interval",
      cex.main = 1.5,   font.main= 4, col.main= "blue",
      cex.sub = 1, font.sub = 3, col.sub = "red")
box()

plot(factor(year2011),xlim=c(0,6),ylim=c(0,40))
title(main="b)",sub="Sample size 15", ylab="Frequency",xlab="Calving interval",col.main=4,cex.main = 1.5,   font.main= 4, col.main= "blue",
      cex.sub = 1, font.sub = 3, col.sub = "red")
box()

plot(factor(year2012),xlim=c(0,6),ylim=c(0,40))
title(main="c)",sub="Sample size 25", ylab="Frequency",xlab="Calving interval",col.main=4,cex.main = 1.5,   font.main= 4, col.main= "blue",
      cex.sub = 1, font.sub = 3, col.sub = "red")
box()

plot(factor(year2013),xlim=c(0,6),ylim=c(0,40))
title(main="c)",sub="Sample size 45", ylab="Frequency",xlab="Calving interval",col.main=4,cex.main = 1.5,   font.main= 4, col.main= "blue",
      cex.sub = 1, font.sub = 3, col.sub = "red")
box()

```
 
What is interesting is that in 2011 there were more intervals that subsequently reduced the standard error of the estimate and the precision increased. 


**My idea of why this has happened is as follows:**

At the time of collection in 2011 the number of possible intervals greater than 5 was very unlikely (as there where only 5 years of research). This meant that the overall error of the estimate was reduced as there was very few "chances" of obtaining a calving interval of 5 or 6 even though there may have been quite a few. The following estimates then picked these up and the error in the estimate increased again. 
 
 
```{r fancy plotting skills, include=FALSE}
theme_Publication <- function(base_size=14, base_family="helvetica") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "bottom",
               legend.direction = "horizontal",
               legend.key.size= unit(0.2, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}
```

The upgrade for publication standard:
**working in the R enviroment**

```{r raw graph 3, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
library(qpcR)
#data in one way for plot
rawdata <- qpcR:::cbind.na(year2010,year2011,year2012,year2013)
rawdata <- as.data.frame(rawdata)

#in correct format for ggplot2
year2010 <- data.frame(year2010,year = c("2010"))
year2010 <- rename(year2010, interval = year2010, year = year )
year2011 <- data.frame(year2011,year = c("2011"))
year2011 <- rename(year2011, interval = year2011, year = year )
year2012 <- data.frame(year2012,year = c("2012"))
year2012 <- rename(year2012, interval = year2012, year = year )
year2013 <- data.frame(year2013,year = c("2013"))
year2013 <- rename(year2013, interval = year2013, year = year )
ggplotraw <- rbind(year2010,year2011,year2012, year2013)
ggplotraw$interval <- as.numeric(as.character(ggplotraw$interval))

#PLOTS
#1
# ggplot(ggplotraw,aes(x = interval))+
#   geom_bar(width = 0.5)+
#   facet_wrap(~year)

#I have found the problem of the previous version I think it is with theme_hc ...but I have not found a solution yet
#2

# ggplot(year2013,aes(x = interval)) +
#     geom_bar(alpha = 1, width = 0.9,fill = "black") +
#     xlab(label = "Calving interval") +
#     #geom_density(adjust = 0.1,alpha=.2, fill="#FF6666")
#     ylab(label = "Total number of observation") +
#     scale_y_continuous(breaks = c(0,5,10,15,20,25,28)) +
#     theme(panel.grid.minor = element_line(colour="white", size=0.5)) +
#     theme_classic()
  

# ggplot(year2013,aes(x = interval), ylim = c(0,35)) +
#     geom_bar(alpha = 1, width = 0.9,fill = "black") +
#     xlab(label = "Calving interval") +
#     #geom_density(adjust = 0.1,alpha=.2, fill="#FF6666")
#     ylab(label = "Total number of observation") +
#     scale_y_continuous(breaks = c(0,5,10,15,20,25,28)) +
#     theme_classic()+
#   theme(panel.grid.minor = element_line(colour="white", size=0.5)) +
#     theme(panel.border = element_rect(fill=NA, colour = "black", size=2)) 

######final after comments
#observation"S"
#up to 30
# dev.off()
# ggplot(year2013,aes(x = interval)) +
#     geom_bar(alpha = 1, width = 0.9,fill = "black") +
#     xlab(label = "Calving interval") +
#     #geom_density(adjust = 0.1,alpha=.2, fill="#FF6666")
#     ylab(label = "Total number of observations") +
#     scale_y_continuous(breaks = c(0,5,10,15,20,25,30), limits = c(0,30)) +
#    # ylim (0,32) +
#     theme(axis.line = element_line(colour = 'black', size = 1),
#           axis.ticks = element_line(colour = "black", size = 1),
#           panel.border = element_blank(), 
#           panel.grid.major = element_blank(), 
#           panel.grid.minor = element_blank(), 
#           legend.key = element_blank(), 
#           strip.background = element_rect(fill = "white", colour = "black", size = 1),
#           panel.background = element_rect(fill = "white", 
#             colour = NA),
#           axis.text = element_text(size = rel(0.8), 
#             colour = "black"))
# 
# theme_classic()
# theme_bw
# theme_grey

png("Figure_2_NZSRW_calving_interval_2017_highres.png", width = 12, height = 14.8, units = 'cm', res = 1200)

ggplot(year2013,aes(x = interval)) +
    geom_bar(alpha = 1, width = 0.9,fill = "black") +
    xlab(expression("Calving"~"interval"~(italic("years")))) +
    ylab(expression("Total"~"number"~"of"~"observations"~(italic("n")))) +
    scale_y_continuous(breaks = c(0,5,10,15,20,25,30), limits = c(0,30)) +
    theme(axis.line = element_line(colour = 'black', size = 0.65),
          axis.ticks = element_line(colour = "black", size = 0.65),
          panel.border = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          legend.key = element_blank(), 
          strip.background = element_rect(fill = "white", colour = "black", size = 1),
          panel.background = element_rect(fill = "white", 
            colour = NA),
          axis.text = element_text(size = rel(0.8), 
            colour = "black"))

dev.off()


png("Figure_2_NZSRW_calving_interval_2017_lowres.png", width = 12, height = 14.8, units = 'cm', res = 400)

ggplot(year2013,aes(x = interval)) +
    geom_bar(alpha = 1, width = 0.9,fill = "black") +
    xlab(expression("Calving"~"interval"~(italic("years")))) +
    ylab(expression("Total"~"number"~"of"~"observations"~(italic("n")))) +
    scale_y_continuous(breaks = c(0,5,10,15,20,25,30), limits = c(0,30)) +
    theme(axis.line = element_line(colour = 'black', size = 0.65),
          axis.ticks = element_line(colour = "black", size = 0.65),
          panel.border = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          legend.key = element_blank(), 
          strip.background = element_rect(fill = "white", colour = "black", size = 1),
          panel.background = element_rect(fill = "white", 
            colour = NA),
          axis.text = element_text(size = rel(0.8), 
            colour = "black"))

dev.off()


#inculdes arrows but looks shit

ggplot(year2013,aes(x = interval)) +
    geom_bar(alpha = 1, width = 0.9,fill = "black") +
    xlab(label = "Calving interval (years)") +
    #geom_density(adjust = 0.1,alpha=.2, fill="#FF6666")
    ylab(label = "Total number of observations") +
    scale_y_continuous(breaks = c(0,5,10,15,20,25,30), limits = c(0,30)) +
   # ylim (0,32) +
    theme(axis.line = element_line(colour = 'black', size = 1),
          axis.line.x = element_line(arrow = arrow(length = unit(4, 'mm'))), 
    axis.line.y=element_line(arrow = arrow(length = unit(4, 'mm'))),
          axis.ticks = element_line(colour = "black", size = 1),
          panel.border = element_blank(), 
          panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(), 
          legend.key = element_blank(), 
          strip.background = element_rect(fill = "white", colour = "black", size = 1),
          panel.background = element_rect(fill = "white", 
            colour = NA),
          axis.text = element_text(size = rel(0.8), 
            colour = "black"))
         # element_grob.element_line())
     # annotate("segment", x=-Inf,xend=Inf,y=0,yend=0,arrow=arrow())



```

<!-------------------------------------------------------------------------->
#Missing intervals (Bradford et al. 2008)

This is a way to see what the calving interval might be if we had in fact missed calving events that happened before or after the study period.

```{r missing intervals, echo=FALSE, fig.height=10, message=FALSE, warning=FALSE}
#################################Missing calving intervals################
#Intervals modified by accounting for missed intervals
#Bradford et al. 2008

#Raw Data
RealCI <- as.numeric(year2013$interval)

#Confidence interval
xlong <- RealCI
meanlong<-sum(xlong)/length(xlong)
slong<-sd(xlong)
SElong<-slong/(sqrt(length(xlong)))
nlong<-(length(xlong))
#Standard error and confidence intervals
#2 sided t value at the 95% level = 2.093
lowqtlong <- meanlong-(qt(0.975,nlong)*SElong)
highqtlong <- meanlong+(qt(0.975,nlong)*SElong)

####################MED CI########################################
# 2x 6's and 1x 5 replaced with 3threes
MedCI <- c(RealCI[RealCI < 5],3,3,3,3,2,3)  
#sort(MedCI)
xmed<-MedCI
meanmed<-sum(xmed)/length(xmed)
smed<-sd(xmed)
SEmed<-smed/(sqrt(length(xmed)))
nmed<-(length(xmed))

#Standard error and confidence intervals
lowqtmed <- meanmed-(qt(0.975,length(xmed))*SEmed)
highqtmed <- meanmed+(qt(0.975,length(xmed))*SEmed)


############################SHORT CI##################################
#6,5 replaced with 2 year intervals

LowCI <- c(RealCI[RealCI < 4],3,3,3,3,3,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2)  
xshort<-LowCI
meanshort<-mean(xshort)
sshort<-sd(xshort)
SEshort<-sshort/(sqrt(length(xshort)))

#Standard error and confidence intervals
lowqtshort <- meanshort-(qt(0.975,length(xshort))*SEshort)
highqtshort <- meanshort+(qt(0.975,length(xshort))*SEshort)

bdata <-qpcR:::cbind.na(RealCI,MedCI,LowCI)
bdata <- as.data.frame(bdata)

#Structure of data set
#str(bdata)
```

```{r missing intervals plot, fig.height=5.5, fig.width=4.5, message=FALSE, warning=FALSE, include=FALSE}
#Basic plots 
par(mfrow=c(3,1))
plot(factor(bdata$LowCI),main="Lowest possible interval")
plot(factor(bdata$MedCI), main="Medium possible interval")
plot(factor(bdata$RealCI),main="Observed interval")
```

```{r missing intervals plot2, fig.height=5.5, fig.width=4.5, message=FALSE, warning=FALSE, include=FALSE}
#Density basic plots 
par(mfrow=c(3,1))
plot(density(as.numeric(as.character(LowCI)),bw=.5), main="Lowest possible interval")
plot(density(as.numeric(as.character(MedCI)),bw= 0.5), main="Medium possible interval")
plot(density(as.numeric(as.character(RealCI)),bw = 0.5),main="Observed interval")
```

```{r missing intervals table, fig.height=8, message=FALSE, warning=FALSE, include=FALSE}

###################################SUMMARY############################
#Pull out important information
Sumtable<-data.frame(variable = c("low.qt","mean","high.qt","sd", "SE"),                 short=c(lowqtshort,meanshort,highqtshort,sshort,SEshort),
                     medium=c(lowqtmed,meanmed,highqtmed,smed,SEmed),
                     real=c(lowqtlong,meanlong,highqtlong,slong,SElong))

#Make dataframe to plot
n <- c(length(LowCI),length(MedCI),length(year2013$interval))
mY <- c(mean(LowCI),mean(MedCI),mean(year2013$interval))
interval <-c("Low", "Medium","Observed")
low.qt <- c(lowqtshort,lowqtmed,low.qt2013)
high.qt <- c(highqtshort,highqtmed,high.qt2013)
sd <- c(sshort,smed,s2013)
Sumtable <- cbind(interval,n,mY,low.qt,high.qt,sd)
Sumtable <-  as.data.frame(Sumtable)

 Sumtable$n <- as.numeric(as.character(Sumtable$n))
 Sumtable$mY <- as.numeric(as.character(Sumtable$mY))
 Sumtable$low.qt <- as.numeric(as.character(Sumtable$low.qt))
 Sumtable$high.qt <- as.numeric(as.character(Sumtable$high.qt))
 Sumtable$sd <- as.numeric(as.character(Sumtable$sd))
 Sumtable$interval <- as.character(Sumtable$interval)
```

```{r missing intervals plot3, echo=FALSE, fig.height=4, message=FALSE, warning=FALSE}
library(ggthemes)
library(ggplot2)
library(gridExtra)

# #different graph options
# ggplot(Sumtable, aes(y = mY, x = interval,color = interval)) +
#   geom_point(size = 3) +
#   geom_errorbar(aes(ymin = low.qt, ymax = high.qt), width = 0.3,size = 1.1) +
#   scale_y_continuous(breaks = round(seq(2.3, 3.6, by = 0.2),1)) +
#   labs(y = "Mean Calving Interval",x = "Calving interval modification" ) +
#   #axis(side = 1,outer = ) +
#   theme(legend.position = "none", axis.text.x = element_text(size=12),    axis.text.y = element_text(size=12))
# 
# ###2
# ggplot(Sumtable, aes(y = mY, x = interval,color = interval)) +
#   geom_errorbar(aes(ymin = low.qt, ymax = high.qt), width = 0.3,size = 1.1) +
#   scale_y_continuous(breaks = round(seq(2.3, 3.6, by = 0.1),1)) +
#   labs(y = "Mean Calving Interval",x = "Calving interval modification" ) +
#   geom_point(size = 3) +
#   theme_hc() +
#   theme(legend.position="none")
# 
# 
# ###3
# ggplot(Sumtable, aes(y = mY, x = interval,color = interval)) +
#   geom_errorbar(aes(ymin = low.qt, ymax = high.qt), width = 0.3,size = 1.1) +
#   scale_y_continuous(breaks = round(seq(2.3, 3.6, by = 0.1),1)) +
#   labs(y = "Mean Calving Interval",x = "Calving interval modification" ) +
#   geom_point(size = 3) +
#   theme_gdocs()+
#   theme(legend.position="none")
# 
# 
# ####4
# ggplot(Sumtable, aes(y = mY, x = interval,color = interval)) +
#   geom_errorbar(aes(ymin = low.qt, ymax = high.qt), width = 0.3,size = 1.1) +
#   scale_y_continuous(breaks = round(seq(2.3, 3.6, by = 0.1),1)) +
#   labs(y = "Mean Calving Interval",x = "Calving interval modification" ) +
#   geom_point(size = 3) +
#   theme_classic() +
#   theme_hc() +
#   theme(legend.position="none")


####5
ggplot(Sumtable, aes(y = mY, x = interval)) +
  geom_errorbar(aes(ymin = low.qt, ymax = high.qt), width = 0.3,size = 1.1) +
  scale_y_continuous(breaks = round(seq(2.3, 3.6, by = 0.2),1)) +
  labs(y = "Mean calving interval",x = "Calving interval modification" ) +
  geom_point(size = 3) +
  theme_classic() +
  theme_hc() +
  theme(legend.position="none")
  
```


The error gets smaller as the variance in the estimate decreases. We are artificially doing this here by modifying data to match what is more biologically possible.


```{r missing_data_table, echo=FALSE}
library(knitr)

kable(Sumtable, format = "markdown")

```


<!-------------------------------------------------------------------->
#Bootstrapping

**Table of parameters:** This is taken directly from my Master's thesis. There may be updated estimates that need to be checked here.

```{r srw_data_table, echo=FALSE}
library(knitr)
srwdat <- read.csv(file = "srw_data.csv")

#str(srwdat)
kable(srwdat, format = "markdown")

```

Bootstrap mean calving interval 1000 times and save the mean for each bootstrap sample. Here I investigate the effect of a sample size of 10,100,1000,2000 from the observed NZ calving interval.

<!--###The plot below shows a single bootstrap sample for each of the sample sizes:-->

```{r bootstrap single, fig.height=5, include=FALSE}
############################NZ Simple sample##############################
#WITH replacement

# to try and match number of intervals observed in other populations
# find references
SAreps <- 1500 
ARreps <- 800
Aussiereps <- 2000
low <- 1000
verylow <- 100
lowest <- 10

#Very raw plots
par(mfrow=c(2,3))
plot(factor(sample(year2013$interval,lowest,replace=T)),main = "3 intervals")
plot(factor(sample(year2013$interval,verylow,replace=T)),main = "10 intervals")
plot(factor(sample(year2013$interval,low,replace=T)),main = "30 intervals")
plot(factor(sample(year2013$interval,Aussiereps,replace=T)),main = "500 intervals")
plot(factor(sample(year2013$interval,ARreps,replace=T)),main = "800 intervals")
plot(factor(sample(year2013$interval,SAreps,replace=T)),main = "1500 intervals")
```


```{r bootstrap_multiple, include=FALSE}
#do each one 1000 times
boots <- 1000
n <- c(1:1000)


###########################n10
var10 <- paste0("n_", 1:10)
sample10 <-matrix(data = NA, ncol = lowest, nrow = boots)
colnames(sample10) <- as.list(var10)

for (i in 1:boots) {
                    sample10 [i, ] <- sample(year2013$interval,lowest,replace=T)
                        }  #i

sample10 <- as.data.frame(sample10)
sample10 <- sample10 %>%
            mutate(mean10 = rowMeans(sample10))

sample10t <- as.matrix(sample10)
sample10t <-t(sample10t)

#########################verylow sample size
#set up variable names
var100 <- paste0("n_", 1:100)

sample100 <-matrix(data = NA, ncol = verylow, nrow = boots)
colnames(sample100) <- as.list(var100)

for (i in 1:boots) {
                    sample100 [i, ] <- sample(year2013$interval,verylow,replace=T)
                        }  #i

sample100 <- as.data.frame(sample100)
sample100 <- sample100 %>%
            mutate(mean100 = rowMeans(sample100))

#########################middle one
#set up variable names
var500 <- paste0("n_", 1:500)

sample500 <-matrix(data = NA, ncol = 500, nrow = boots)
colnames(sample500) <- as.list(var500)

for (i in 1:boots) {
                    sample500 [i, ] <- sample(year2013$interval,500,replace=T)
                        }  #i

sample500 <- as.data.frame(sample500)
sample500 <- sample500 %>%
            mutate(mean500 = rowMeans(sample500))


#########################low sample size
#set up variable names
var1000 <- paste0("n_", 1:1000)

sample1000 <-matrix(data = NA, ncol = low, nrow = boots)
colnames(sample1000) <- as.list(var1000)

for (i in 1:boots) {
                    sample1000 [i, ] <- sample(year2013$interval,low,replace=T)
                        }  #i

sample1000 <- as.data.frame(sample1000)
sample1000 <- sample1000 %>%
            mutate(mean1000 = rowMeans(sample1000))

#########################AUS sample size
#set up variable names
varA <- paste0("n_", 1:2000)

sampleA <-matrix(data = NA, ncol = Aussiereps, nrow =  boots)
colnames(sampleA) <- as.list(varA)

for (i in 1:boots) {
                    sampleA [i, ] <- sample(year2013$interval,Aussiereps,replace=T)
                        }  #i

sampleA <- as.data.frame(sampleA)
sampleA <- sampleA %>%
            mutate(meanA = rowMeans(sampleA))

sampleAt <- t(sampleA)

for(i in c(1:ncol(sampleA))) {
    sampleA[,i] <- as.numeric(as.character(sampleA[,i]))
}




#COnfidence intervals

ab <- sort(sampleA$meanA)
nab <- length(ab)
#low = 25/1000
ab2.5 <- ab[25]
#high = 975/1000
ab0.97.5 <- ab[975]

ab <- sort(sampleA$meanA)
nab <- length(ab)
#low = 25/1000
ab2.5 <- ab[25]
#high = 975/1000
ab0.97.5 <- ab[975]

```

###Publication plot
1000 replicates
```{r bootstrap plot2, echo=FALSE, fig.height=5, message=FALSE, warning=FALSE}
#plot the data over each other to look at change in density
par(mfrow=c(1,1))
#plot(density(sample3$mean3,bw = .15),lwd = 3,lyt = 5, main = "", xlab = "Calving interval", box = FALSE,axis = FALSE)

plot(density(sample10$mean10,bw = .05),col ="black", lty = 1, main = "", lwd = 5,ylim = c(0,8),xlim = c(2,4.5), axes=FALSE,xlab = "Calving interval")
lines(density(sample100$mean100,bw = .05),col ="black", lty = 2, lwd = 4)
lines(density(sample500$mean500,bw = .05),col ="black", lty = 3, lwd = 3)
lines(density(sample1000$mean1000,bw = .05),col ="black", lty = 4, lwd = 2)
lines(density(sampleA$meanA,bw = .05),col ="black", lty = 5, lwd = 1)
legend('topright',title = "Legend", c("n=10, cv=8.12 ", "n=100, cv=2.43", "n=500, c.v=1.15", "n=1000, cv=0.79", "n=2000, cv=0.56"),bty = "n",
  lty = c(1,2,3,4,5), lwd = c(5,4,3,2,1), cex=.75)
axis(1,lwd=2) 
axis(2,lwd=2)




Sumtable<-data.frame(variable = c("low.qt","mean","high.qt","sd", "SE"),                 short=c(lowqtshort,meanshort,highqtshort,sshort,SEshort),
                     medium=c(lowqtmed,meanmed,highqtmed,smed,SEmed),
                     real=c(lowqtlong,meanlong,highqtlong,slong,SElong))
#try it in ggplot2

x <- data.frame(n10 = sample10$mean10, n100 = sample100$mean100,n500 = sample500$mean500, n1000 = sample1000$mean1000, n2000 = sampleA$meanA)
#str(x)
library(reshape2)
library(ggplot2)
x <- melt(x)
x <- rename(x,sample = variable, smean = value)

```

```{r bootstrap plot3, include=FALSE}
#1
ggplot(x,aes(x = smean, fill = sample)) +
  geom_density(alpha = 0.50) +
  theme_classic() +
  facet_wrap(~sample) +
  theme_hc() +
  theme(legend.position = "none") +
  theme(legend.background = element_rect(fill="lightblue", size=0.5, linetype = "solid", colour = "darkblue"))

# #2
# ggplot(x,aes(x = smean, fill = sample)) +
#   geom_density(alpha = 0.50) +
#   theme_classic() +
#   facet_wrap(~sample) +
#   theme(legend.position = "none")
# 
# #3
# ggplot(x,aes(x = smean, fill = "black", colour = "black")) +
#   geom_density() +
#   theme_classic() +
#   facet_wrap(~sample) +
#   theme(legend.position = "none")
# 
# #4
# samplesize <- c("Sample size 10","Sample size 100","Sample size 1000","Sample size 2000")
# 
# ggplot(x,aes(x = smean, fill = sample)) +
#   geom_density(alpha = 0.50) +
#   scale_x_continuous(name = "Mean ozone in\nparts per billion",
#                               limits=c(2,4 )) +
#   scale_y_continuous(name = "Density",
#                            breaks = seq(0, 5, 20)) +
#   theme_classic() +
#   facet_wrap(~sample)
# 
# #5
# ggplot(x,aes(x = smean, fill = sample)) +
#   geom_density(alpha = 0.50) +
#   scale_x_continuous(name = "Mean calving interval") +
#   scale_y_continuous(name = "Density") +
#   theme_classic() +
#   facet_wrap(~sample) +
#   theme_hc() +
#   theme(legend.position = "none")
# 
# 
# 
#6
#Low = less variation
#High = More variation
CV <- function(mean, sd){
      (sd/mean)*100
}

###########################UP TO HERE
#Sorint this NOW
CVn10 <- CV(mean(sample10$mean10),sd(sample10$mean10))
CVn100 <- CV(mean(sample100$mean100),sd(sample100$mean100))
CVn500 <- CV(mean(sample500$mean500),sd(sample500$mean500))
CVn1000 <- CV(mean(sample1000$mean1000),sd(sample1000$mean1000))
CVnA <- CV(mean(sampleA$meanA),sd(sampleA$meanA))
CVs <- c(CVn10,CVn100,CVn500,CVn1000,CVnA)
CVs <- data.frame(CV = round(CVs,2))


# 
# ggplot(x,aes(x = smean, fill = sample)) +
#   geom_density(alpha = 0.50) +
#   scale_x_continuous(name = "Mean calving interval") +
#   scale_y_continuous(name = "Density") +
#   theme_classic() +
#   facet_wrap(~sample,labeller=label_both) +
# #  geom_text(data=CVs, aes(x=1.2, y=1.2), colour="black") +
#   theme_hc() +
#   theme(legend.position = "none")

```

```{r bootstrap plot4, include=FALSE}
# #1
# ggplot(x,aes(x = smean, fill = sample)) +
#   geom_density(alpha = 0.50) +
#   theme_classic() +
#   theme_hc() +
#   theme(legend.position = c(0.8,0.8)) +
#   theme(legend.background = element_rect(fill="lightblue", size=0.5, linetype = "solid", colour = "darkblue"))
# 
# #2 
ggplot(x,aes(x = smean, fill = sample)) +
  geom_density(alpha = 1) +
  scale_x_continuous(name = "Mean calving interval") +
  scale_y_continuous(name = "Density") +
  theme_classic() +
  theme_hc() +
  theme(legend.position = c(0.8,0.8)) +
  theme(legend.background = element_rect(fill="lightblue", size=0.5, linetype = "solid", colour = "darkblue"))
```



```{r density of bootstrap plot, include=FALSE}
#3
ggplot(x,aes(x = smean, group = sample)) +
  geom_density(alpha = 0.5, fill="gray20") +
  scale_x_continuous(name = "Mean calving interval") +
  scale_y_continuous(name = "Density") +
  theme_classic() 
  

library(ggplot2)
df <- data.frame(cond = factor( rep(c("A","B"), each=200) ), rating = c(rnorm(200),rnorm(200, mean=.8)))
ggplot(df, aes(x=rating, fill=cond)) + geom_density(alpha=.3)+xlab("NEW RATING TITLE")+ylab("NEW DENSITY TITLE")


df <- data.frame(x=1:10,group=c(rep("a",5),rep("b",5)))

legend_title <- "OMG My Title"

ggplot(df, aes(x=x, fill=group)) + geom_density(alpha=.3) +   
    scale_fill_manual(legend_title,values=c("orange","red"))
```



```{r density of bootstrap plot2, echo=FALSE}
#plot the data over each other to look at change in density
par(mfrow=c(1,1))
#plot(density(sample3$mean3,bw = .15),lwd = 3,lyt = 5, main = "", xlab = "Calving interval", box = FALSE,axis = FALSE)

plot(density(sample10$mean10,bw = .05),col ="black", lty = 3, main = "", lwd = 1,ylim = c(0,8),xlim = c(2.5,4.5), axes=FALSE,xlab = "Calving interval")
lines(density(sample100$mean100,bw = .05),col ="black", lty = 4, lwd = 1)
lines(density(sample500$mean500,bw = .05),col ="black", lty = 5, lwd = 1)
lines(density(sample1000$mean1000,bw = .05),col ="black", lty = 2, lwd = 1)
lines(density(sampleA$meanA,bw = .05),col ="black", lty = 1, lwd = 1)
legend('topright',title = "Legend", c("n=10", "n=100", "n=500", "n=1000", "n=2000"),bty = "n",
  lty = c(3,4,5,2,1), lwd = c(1,1,1,1,1), cex=.75)
axis(1,lwd=2) 
axis(2,lwd=2)



#number 2
plot(density(sample10$mean10,bw = .05),col ="black", lty = 3, main = "", lwd = 1,ylim = c(0,8),xlim = c(2.5,4.5), axes=FALSE,xlab = "Calving interval")
lines(density(sample100$mean100,bw = .05),col ="black", lty = 2, lwd = 1)
lines(density(sample500$mean500,bw = .05),col ="black", lty = 1, lwd = 1)
legend('topright',title = "Legend", c("n=10", "n=100", "n=500"),bty = "n",
  lty = c(1,2,3,4,5), lwd = c(1,1,1,1,1), cex=.75)
axis(1,lwd=2) 
axis(2,lwd=2)



#number 3
plot(density(sample10$mean10,bw = .05),col ="black", lty = 3, main = "", lwd = 1,ylim = c(0,8),xlim = c(2.5,4.5), axes=FALSE,xlab = "Calving interval")
lines(density(sample100$mean100,bw = .05),col ="black", lty = 4, lwd = 1)
lines(density(sample500$mean500,bw = .05),col ="black", lty = 2, lwd = 1)
lines(density(sampleA$meanA,bw = .05),col ="black", lty = 1, lwd = 1)
legend('topright',title = "Legend", c("n=10", "n=100", "n=500","n=2000"),bty = "n",
  lty = c(3,2,1,1), lwd = c(1,1,1,1,1), cex=.75)
axis(1,lwd=2) 
axis(2,lwd=2)
```

```{r}

#final [plot]
#size defined by NZJFMR
#  195 mm (h) × 148 mm (w).
#ylab(expression("Total"~"number"~"of"~"observations"~(italic("n")))) +

png("Figure_3_NZSRW_calving_interval_2017_lowres.png", width = 14.8, height = 14.8, units = 'cm', res = 400)

plot(density(sample10$mean10,bw = .05),col ="black", lty = 3, main = "", lwd = 1,ylim = c(0,8),xlim = c(2.5,4.5), axes=FALSE, xlab = expression("Calving"~"interval"~(italic("years"))))
lines(density(sample100$mean100,bw = .05),col ="black", lty = 4, lwd = 1)
lines(density(sample500$mean500,bw = .05),col ="black", lty = 5, lwd = 1)
lines(density(sample1000$mean1000,bw = .05),col ="black", lty = 2, lwd = 1)
lines(density(sampleA$meanA,bw = .05),col ="black", lty = 1, lwd = 2)
legend(y = 8, x = 3.9,title = expression(bold("Sample size (n)")), c(expression(italic("n")~"="~"10"), expression(italic("n")~"="~"100"), expression(italic("n")~"="~"500"), expression(italic("n")~"="~"1000"), expression(italic("n")~"="~"2000")),bty = "n",
  lty = c(3,4,5,2,1), lwd = c(1,1,1,1,2), cex=1) 
 axis(1,lwd=2)
axis(2,lwd=2)
dev.off()


png("Figure_3_NZSRW_calving_interval_2017_highres.png", width = 14.8, height = 14.8, units = 'cm', res = 1200)

plot(density(sample10$mean10,bw = .05),col ="black", lty = 3, main = "", lwd = 1,ylim = c(0,8),xlim = c(2.5,4.5), axes=FALSE,xlab = expression("Calving"~"interval"~(italic("years"))))
lines(density(sample100$mean100,bw = .05),col ="black", lty = 4, lwd = 1)
lines(density(sample500$mean500,bw = .05),col ="black", lty = 2, lwd = 1)
lines(density(sample1000$mean1000,bw = .05),col ="black", lty = 5, lwd = 1)
lines(density(sampleA$meanA,bw = .05),col ="black", lty = 1, lwd = 2)
legend(y = 8, x = 3.9,title = expression(bold("Sample size (n)")), c(expression(italic("n")~"="~"10"), expression(italic("n")~"="~"100"), expression(italic("n")~"="~"500"), expression(italic("n")~"="~"1000"), expression(italic("n")~"="~"2000")),bty = "n",
  lty = c(3,4,2,5,1), lwd = c(1,1,1,1,2), cex=1) 
axis(1,lwd=2)
axis(2,lwd=2)
dev.off()

# png("Figure_3_NZSRW_calving_interval_2017_highres_col.png", width = 14.8, height = 14.8, units = 'cm', res = 1200)
# plot(density(sample10$mean10,bw = .05),col ="black", lty = 3, main = "", lwd = 2,ylim = c(0,8),xlim = c(2.5,4.5), axes=FALSE,xlab = "Calving interval")
# lines(density(sample100$mean100,bw = .05),col ="red", lty = 4, lwd = 2)
# lines(density(sample500$mean500,bw = .05),col ="orange", lty = 5, lwd = 2)
# lines(density(sample1000$mean1000,bw = .05),col ="darkolivegreen", lty = 2, lwd = 2)
# lines(density(sampleA$meanA,bw = .05),col ="royalblue", lty = 1, lwd = 2)
# legend(y = 8, x = 3.9,title = expression(bold("Sample size (n)")), c("n=10", "n=100","n=500","n=1000","n=2000"),bty = "n",
#   lty = c(3,4,5,2,1), lwd = c(2,2,2,2,2), col = c("black","red","orange","darkolivegreen","royalblue"),cex=1) 
# axis(1,lwd=2)
# axis(2,lwd=2)
# dev.off()
# 
# png("Figure_3_NZSRW_calving_interval_2017_highres_col2.png", width = 14.8, height = 14.8, units = 'cm', res = 1200)
# plot(density(sample10$mean10,bw = .05),col ="black", main = "", lwd = 2,ylim = c(0,8),xlim = c(2.5,4.5), axes=FALSE,xlab = "Calving interval")
# lines(density(sample100$mean100,bw = .05),col ="red",  lwd = 2)
# lines(density(sample500$mean500,bw = .05),col ="orange", lwd = 2)
# lines(density(sample1000$mean1000,bw = .05),col ="darkolivegreen", lwd = 2)
# lines(density(sampleA$meanA,bw = .05),col ="royalblue", lwd = 2)
# legend(y = 8, x = 3.9,title = expression(bold("Sample size (n)")), c("n=10", "n=100","n=500","n=1000","n=2000"),bty = "n",
#   lwd = c(2,2,2,2,2), col = c("black","red","orange","darkolivegreen","royalblue"),cex=1) 
# axis(1,lwd=2)
# axis(2,lwd=2)
# dev.off()

```

 

